Restrictive Clearance Comparison

### 2025-02-07 data-raw/RestrictiveClearance_comparison.Rmd
Added clustering chunk and began re-writing analysis steps alongside the respective chunk.

this_pk <- do_prefit(this_pk)
this_pk <- do_fit(this_pk, n_cores = 14)
```

@@

Steps in the analysis:

1. Get chemicals + species combinations in CvTdb that have _httk_ predictions.
2. Safely get parameters, clearance, and solutions for each of these groups.
3. Compare clearances from _httk_ with those from _invivoPKfit_.
4. Compare $f_{up}$ for subset of data with total values available via _httk_.
5. Use new _clearance-aware_ optimization of 1-compartment parameters to
benchmark _httk_ solutions.
6. Estimate the fraction unbound in plasma for chemicals with rat data,
for those same chemicals with human data for Clint and Fup. 
Begin with the assumption that there is restrictive clearance, those chemicals
that do not change when optimized (and produce reasonable fits) are likely to
be _restrictively cleared_, whereas those that increase Fup during maximum
likelihood estimation are _possibly non-restrictively cleared_. 
7. Other factors such as renal clearance may play a role in those cases.
Validate the clearance estimate by measuring changes in predicted clearance.
Important to identify whether there seems to be a "fraction missing".
8. Wrap the previous steps into attempts to identify true positive
categorizations for Restrictive and Potentially Non-Restrictively metabolized
chemicals (include Lipinski's Rule of 5 too).
9. Implement a _slow_ PBPK model that optimizes $f_{up}$ and $CL_{int}$ under
restrictive and non-restrictive clearance conditions. Test whether these have a
substantial difference from initial values. Use `coef_sd()` to evaluate
uncertainty of these parameter estimates for a larger share of chemicals.
10. Compare the results of 9 with the putative/predicted clearance from 1-6.

@@

ggsave(filename = paste0(Sys.getenv("FIG_DIR"),
                         "RestrictiveClearance_Analysis/",
                         "CURRENT_", "Alluvium_Lipinski.png"),
       bg = "white",
       height = 4, width = 6
)
```

@@

ggsave(filename = paste0(Sys.getenv("FIG_DIR"),
                         "RestrictiveClearance_Analysis/",
                         "CURRENT_", "RMSLE_comparison_fullChart.png"),
       bg = "white",
       height = 4, width = 8
)

@@

@@

Now I will try to triangulate the model weights witht the Css estimates.

@@

```{r CLtot_subset}

this_tkstats %>%
  inner_join(this_httk_clearance_df) %>%
  inner_join(combined_model_weights) %>%
  filter(!is.na(CLtot.tkstats)) %>%
  mutate(comparison_CLtot = factor(
    case_when(
    between(log2(CLtot.tkstats), 
            log2(restrictive_clearance)-1,
            log2(restrictive_clearance)+1) ~ "Restrictive",
    between(log2(CLtot.tkstats), 
            log2(nonrestrictive_clearance)-1,
            log2(nonrestrictive_clearance)+1) ~ "Nonrestrictive",
    between(CLtot.tkstats,
            restrictive_clearance, nonrestrictive_clearance) ~ "Between",
    .default = "Out of Range"
  ), levels = c("Nonrestrictive", "Between", "Restrictive", "Out of Range")
  )) %>%
  ggplot(aes(forcats::fct_reorder(Chemical,
                          gamma_weight))) +
  geom_point(aes(y = CLtot.tkstats, fill = gamma_weight),
             shape = 21) +
  geom_linerange(aes(ymin = restrictive_clearance, 
                     ymax = nonrestrictive_clearance)) +
  geom_point(aes(y = restrictive_clearance),
             shape = 2, size = 0.8) + # Note that this is the lower value
  geom_point(aes(y = nonrestrictive_clearance),
             shape = 6, size = 0.8) +
  scale_fill_viridis_b(option = "viridis", 
                        guide = guide_colorbar(position = "bottom"),
                       limits = c(0,1.5)) +
  scale_y_log10() +
  facet_grid(cols = vars(comparison_CLtot),
             scales = "free_x",
             space = "free_x") +
  labs(y = "Total Clearance", x = "Chemical", fill = "Inverse Confidence") +
  theme_bw() +
  theme(axis.text.x = element_blank(),
        axis.ticks.x = element_blank(),
        plot.subtitle = element_text(hjust = 0.5),
        legend.title = element_text(vjust = 0.8))

@@

this_tkstats %>%
  inner_join(this_httk_clearance_df) %>%
  inner_join(combined_model_weights) %>%
  left_join(this_css_df) %>%
  filter(!is.na(CLtot.tkstats)) %>%
  mutate(comparison_CLtot = factor(
    case_when(
      between(log2(CLtot.tkstats),log2(restrictive_clearance)-1,
              log2(restrictive_clearance)+1) ~ "Restrictive",
      between(log2(CLtot.tkstats),log2(nonrestrictive_clearance)-1,
              log2(nonrestrictive_clearance)+1) ~ "Nonrestrictive",
      between(CLtot.tkstats,
              restrictive_clearance, nonrestrictive_clearance) ~ "Between",
      .default = "Out of range"
  ), levels = c("Nonrestrictive", "Between", "Restrictive", "Out of range"))) %>%
  ggplot(aes(forcats::fct_reorder(Chemical,
                          gamma_weight))) +
  geom_point(aes(y = CLtot.tkstats, fill = gamma_weight),
             shape = 21) +
  geom_linerange(aes(ymin = restrictive_clearance, 
                     ymax = nonrestrictive_clearance)) +
  geom_point(aes(y = restrictive_clearance),
             shape = 4, size = 0.8) + # Note that this is the lower value
  geom_point(aes(y = nonrestrictive_clearance),
             shape = 4, size = 0.8) +
  scale_fill_viridis_b(option = "viridis", 
                        guide = guide_colorbar(position = "bottom")) +
  scale_y_log10() +
  facet_grid(cols = vars(comparison_CLtot),
             scales = "free_x") +
  labs(y = "Total Clearance", x = "Chemical", fill = "Inverse Confidence",
       subtitle = "Estimates within 20% of in vitro modelled predictions") +
  theme_bw() +
  theme(axis.text.x = element_blank(),
        axis.ticks.x = element_blank(),
        plot.subtitle = element_text(hjust = 0.5),
        legend.title = element_text(vjust = 0.8))

@@

>>>>

this_pk <- do_prefit(this_pk)
this_pk <- do_fit(this_pk, n_cores = 12)
```

@@

@@

ggsave(filename = paste0(Sys.getenv("FIG_DIR"),
                         "RestrictiveClearance_Analysis/",
                         "CURRENT_", "Alluvium_Lipinski.png"),
       bg = "white",
       height = 5, width = 6
)
```

@@

ggsave(filename = paste0(Sys.getenv("FIG_DIR"),
                         "RestrictiveClearance_Analysis/",
                         "CURRENT_", "RMSLE_comparison_fullChart.png"),
       bg = "white",
       height = 4.5, width = 8
)

@@

ggsave(filename = paste0(Sys.getenv("FIG_DIR"),
                         "RestrictiveClearance_Analysis/",
                         "CURRENT_", "Css_difference_httk_lineplot.png"),
       bg = "white",
       height = 5, width = 7)

@@

Now I will try to triangulate the model weights with the Css estimates.

@@

```{r CLtot_subset}
enhanced_tkstats <- this_tkstats %>%
  inner_join(this_httk_clearance_df) %>%
  inner_join(combined_model_weights) %>%
  filter(!is.na(CLtot.tkstats)) %>%
  mutate(comparison_CLtot = factor(
    case_when(
    between(log2(CLtot.tkstats), 
            log2(restrictive_clearance)-1,
            log2(restrictive_clearance)+1) ~ "Restrictive",
    between(log2(CLtot.tkstats), 
            log2(nonrestrictive_clearance)-1,
            log2(nonrestrictive_clearance)+1) ~ "Nonrestrictive",
    between(CLtot.tkstats,
            restrictive_clearance, nonrestrictive_clearance) ~ "Between",
    .default = "Out of Range"
  ), levels = c("Nonrestrictive", "Between", "Restrictive", "Out of Range")
  ))
enhanced_tkstats %>%
  ggplot(aes(forcats::fct_reorder(Chemical,
                          gamma_weight))) +
  geom_point(aes(y = CLtot.tkstats, fill = gamma_weight),
             shape = 21) +
  geom_linerange(aes(ymin = restrictive_clearance, 
                     ymax = nonrestrictive_clearance)) +
  geom_point(aes(y = restrictive_clearance),
             shape = 2, size = 0.8) + # Note that this is the lower value
  geom_point(aes(y = nonrestrictive_clearance),
             shape = 6, size = 0.8) +
  scale_fill_viridis_b(option = "viridis", 
                        guide = guide_colorbar(position = "bottom"),
                       limits = c(0,1.5)) +
  scale_y_log10() +
  facet_grid(cols = vars(comparison_CLtot),
             scales = "free_x",
             space = "free_x") +
  labs(y = "Total Clearance", x = "Chemical", fill = "Inverse Confidence") +
  theme_bw() +
  theme(axis.text.x = element_blank(),
        axis.ticks.x = element_blank(),
        plot.subtitle = element_text(hjust = 0.5),
        legend.title = element_text(vjust = 0.8))

@@

this_tkstats %>%
  inner_join(this_httk_clearance_df) %>%
  inner_join(combined_model_weights) %>%
  left_join(this_css_df) %>%
  filter(!is.na(CLtot.tkstats)) %>%
  mutate(comparison_CLtot = factor(
    case_when(
      between(log2(CLtot.tkstats),log2(restrictive_clearance)-1,
              log2(restrictive_clearance)+1) ~ "Restrictive",
      between(log2(CLtot.tkstats),log2(nonrestrictive_clearance)-1,
              log2(nonrestrictive_clearance)+1) ~ "Nonrestrictive",
      between(CLtot.tkstats,
              restrictive_clearance, nonrestrictive_clearance) ~ "Between",
      .default = "Out of range"
  ), levels = c("Nonrestrictive", "Between", "Restrictive", "Out of range"))) %>%
  ggplot(aes(forcats::fct_reorder(Chemical,
                          gamma_weight))) +
  geom_point(aes(y = CLtot.tkstats, fill = gamma_weight),
             shape = 21) +
  geom_linerange(aes(ymin = restrictive_clearance, 
                     ymax = nonrestrictive_clearance)) +
  geom_point(aes(y = restrictive_clearance),
             shape = 4, size = 0.8) + # Note that this is the lower value
  geom_point(aes(y = nonrestrictive_clearance),
             shape = 4, size = 0.8) +
  scale_fill_viridis_b(option = "viridis", 
                        guide = guide_colorbar(position = "bottom")) +
  scale_y_log10() +
  facet_grid(cols = vars(comparison_CLtot),
             scales = "free_x") +
  labs(y = "Total Clearance", x = "Chemical", fill = "Inverse Confidence",
       subtitle = "Estimates within 2-fold of in vitro modelled predictions") +
  theme_bw() +
  theme(axis.text.x = element_blank(),
        axis.ticks.x = element_blank(),
        plot.subtitle = element_text(hjust = 0.5),
        legend.title = element_text(vjust = 0.8))

@@

```{r rudimentary_clustering}

# Set up a data.frame that has chemical, Css difference, Funbound z-score,
# nearest_clearance (using Css), and model weight
this_fupZ_df <- this_param_df %>% select(Chemical, Funbound.plasma.restrictive) %>%
  mutate(Fup_Z = as.vector(scale(sqrt(Funbound.plasma.restrictive))),
         .keep = "unused")

this_dist_prepdf <- this_css_df %>% select(Chemical, model, Route,
                       nearest_clearance, comparison_group_Css,
                       Css_httk_r, Css_httk_nr) %>%
  mutate(relative_difference = log1p(Css_httk_r) - log1p(Css_httk_nr),
         .keep = "unused") %>%
  inner_join(combined_model_weights %>%
               select(Chemical, Species, gamma_weight)) %>%
  inner_join(enhanced_tkstats %>%
               select(Chemical, Species, comparison_CLtot)) %>%
  inner_join(this_fupZ_df) %>%
  distinct() %>%
  mutate(ID = paste(Chemical, Species, model, Route, sep = "."), .keep = "unused")

this_dist_prepdf %>%
ggplot(aes(x = gamma_weight, y = Fup_Z, color = relative_difference)) + 
  geom_point() +
  scale_color_viridis_c() +
  facet_grid(cols = vars(comparison_CLtot))
